JD notes on effects

### 2022-05-25 manuscript/body.tex
Predictions and predictor effects?

Simple OLS regression models are flexible and can easily be generalized, as generalized linear (mixed) models (GL(M)Ms), to examine more complex relationships, including \bmb{nonlinear relationships -- need to clarify}\BC{clarity with BMB} between response and predictors e.g., log transformed response or predictors, interactions between predictors (and via splines for example), and nonlinear transformations via link functions \bmb{due to their flexibility}\BC{clarify why BMB circled it}. This flexibility comes with additional complications, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest \citep{berry_improving_2012, leeper2017interpreting}. Further, coefficient estimates of GLM models involving nonlinear link functions lack direct interpretation\citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model. An alternative is to explore the \emph{effects} of predictors on the predictions of the outcome at various levels of the predictor of interest -- \emph{predictions and effects}\jd{Are we still saying these (prediction and effects) are the same? Or are we now saying they're different, with the same citations?} \citep{fox2009effect, leeper2017package, lenth2018package}. For example, as a way to plan, public health officials may want to know the effects of household income, wealthindex, etc, on the predicted probability that slum dwellers will have improved water services.

@@

We start by generating two predictions from the fitted model \nameref{S1_Appendix}: 1) one with cubic polynomial of \code{age} as a focal predictor; and 2) one with a simple focal predictor (non-polynomial) of \code{Wealthindex}. Then we compared non-anchored (more of predictions) and mean-anchored (effects) confidence intervals around the predicted values (see Fig~\ref{fig:pred_cubic_plots}).

\begin{figure}[h]
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction-styled and effect-styled plots for cubic polynomial and simple focal predictors.} A: The focal predictor is a cubic polynomial. B: The focal predictor is not a polynomial function, implying that the complex interaction is in non-focal predictor. In both cases, the mean-anchored effects are narrower than the non-anchored (predictions), but crosses at the mean of the focal predictor (model center) only in the case of simple (non-polynomial) focal predictor -- B. In the case of cubic polynomial focal predictor, the model center is not a point in the predictor space, i.e., mean of the focal predictor, but a combination of all the model variables in the cubic polynomial, consequently, mean-anchored curves do not cross at a point -- A. The horizontal black and red dotted lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively; the inner red dotted curves and crossing lines are the anchored confidence bands for A and B, respectively; while the outer blue curves are the non-anchored confidence bands in both cases. The central curve and trend line, in A and B, respectively; are the predicted household size based on age and wealth index, respectively.}
\label{fig:pred_cubic_plots}
\end{figure}

In the absence of complex higher order interaction terms or transformations on the focal predictor in simple linear models, we would expect, at the model center, the observed average, $\over{hh~size}$, and the average of the predictions, $\over{\widehat{hh~size}}$, to be identical and cross with the mean-anchored confidence bands (effects), at the model center (see Fig~\ref{fig:pred_cubic_plots}B). On the other hand, if the focal predictor is characterized by complex interactions or any other form of transformation, the model center is not necessarily a point in the predictor space. In this case, the mean-anchored confidence bands will be narrower than the non-anchored ones but not necessarily crossing at the model center (see Fig~\ref{fig:pred_cubic_plots}A).

>>>>

Simple OLS regression models are flexible and can easily be generalized, as generalized linear (mixed) models (GL(M)Ms), to examine more complex relationships, including \bmb{nonlinear relationships -- need to clarify}\BC{clarity with BMB} between response and predictors e.g., log transformed response or predictors, interactions between predictors (and via splines for example), and nonlinear transformations via link functions \bmb{due to their flexibility}\BC{clarify why BMB circled it}. This flexibility comes with additional complications, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest \citep{berry_improving_2012, leeper2017interpreting}. Further, coefficient estimates of GLM models involving nonlinear link functions lack direct interpretation\citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model. An alternative is to explore the \emph{effects} of predictors on the predictions of the outcome at various levels of the predictor of interest -- \emph{predictions and effects}\jd{Are we still saying these (prediction and effects) are the same? Or are we now saying they're different, with the same citations?} \citep{fox2009effect, leeper2017package, lenth2018package}. For example, as a way to plan, public health officials may want to know the effects of household \code{income}, \code{wealthindex}, etc, on the predicted probability that slum dwellers will have improved water services.

@@

We start by generating two predictions from the fitted model \nameref{S1_Appendix}: 1) one with cubic polynomial of \code{age} as a focal predictor; and 2) one with a simple focal predictor (non-polynomial) of \code{Wealthindex}. Then we compared non-anchored and mean-anchored (effects) confidence intervals around the predicted values (see Fig~\ref{fig:pred_cubic_plots}).

\begin{figure}[h]
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction-styled and effect-styled plots for cubic polynomial and simple focal predictors.} A: The focal predictor is a cubic polynomial. B: The focal predictor is a linear function, with a cubic polynomial non-focal predictor. In both cases, the mean-anchored effects are narrower than the non-anchored (predictions), but intersects at the mean of the focal predictor (model center) only in the case of linear focal predictor -- B. In the case of cubic polynomial focal predictor, the model center is not a point in the predictor space, i.e., mean of the focal predictor, but a combination of all the model variables in the cubic polynomial, consequently, mean-anchored curves will not necessarily cross at a point -- A. The horizontal black and red dashed lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively; the inner red dotted curves and crossing lines are the effects for A and B, respectively; while the outer blue curves are the \BC{predictions (place holder for now)} in both cases. The central curve and trend line, in A and B, respectively; are the predicted household size based on age and wealth index, respectively. \bmb{Use different line types ...}\BC{Confirm with BMB}}
\label{fig:pred_cubic_plots}
\end{figure}

In the absence of complex higher order interaction terms or transformations on the focal predictor in simple linear models, we would expect, at the model center, the observed average, $\over{hh~size}$, and the average of the predictions, $\over{\widehat{hh~size}}$, to be identical and all intersect at the mean-anchored confidence bands (effects), at the model center (Fig~\ref{fig:pred_cubic_plots}B). On the other hand, if the focal predictor is characterized by complex interactions or any other form of transformation, the model center is not necessarily a point in the predictor space. In this case, the mean-anchored confidence bands will be narrower than the non-anchored ones but will not necessarily cross at the model center (Fig~\ref{fig:pred_cubic_plots}A).
